MSCA 31006: Final Project - AQI Forecasting

### 2023-08-14 MSCA_31006_Time_Series_Final_Project.Rmd
Add files via upload

```
#### Step 5b: Extract values for time series and plot the series - Monthly

@@

# Forecast the seasonal naïve for (2021-01 to 2023-07)
forecast_snaive_monthly <- snaive(train_monthly, h=31)
# Plot the forecasts for snaive model
plot(forecast_snaive_monthly, main = "PM 2.5 AQI Forecast - SNaive",
         xlab = "Week", ylab = "PM 2.5 AQI")
lines(test_monthly)

# Compare the forecast with the actual test data by calculating MAPE and MSE
# Mean Absolute Percentage Error (MAPE)
mape_snaive <- mean(abs((test_monthly - forecast_snaive_monthly$mean)/test_monthly))
mape_snaive

# Mean Squared Error (MSE)
mse_snaive <- mean((test_monthly - forecast_snaive_monthly$mean)^2)
mse_snaive

@@

>>>>

#### Step 10a: Forecast using Neural Network - Monthly

model_nn_weekly <- nnetar(train_weekly,
                           size=5, 
                           decay=0.5,
                           repeats=1000, 
                           lambda="auto")
summary(model_nn_weekly)
forecast_nn_weekly <- forecast(model_nn_weekly, PI=FALSE, h=110)

# Plot the forecasts for NN model
plot(forecast_nn_weekly, main = "PM 2.5 AQI Forecast - Neural Network",
         xlab = "Weekly", ylab = "PM 2.5 AQI")
lines(test_weekly)

# Compare the forecast with the actual test data by calculating MAPE and MSE
# Mean Absolute Percentage Error (MAPE)
mape_nn_weekly <- mean(abs((test_weekly - forecast_nn_weekly$mean)/test_weekly))
mape_nn_weekly

# Mean Absolute Error (MAE)
mae_nn_weekly <- mean(abs((test_weekly - forecast_nn_weekly$mean)))
mae_nn_weekly

# Mean Squared Error (MSE)
mse_nn_weekly <- mean((test_weekly - forecast_nn_weekly$mean)^2)
mse_nn_weekly

# Evaluate the residuals
checkresiduals(forecast_nn_weekly)
```

#### Step 5b: Extract values for time series and plot the series - Monthly

@@

# Forecast the seasonal naïve for (2021-01 to 2023-07)
forecast_snaive_monthly <- snaive(train_monthly, h=31)
# Plot the forecasts for snaive model
plot(forecast_snaive_monthly, main = "PM 2.5 AQI Forecast - SNaive",
         xlab = "Month", ylab = "PM 2.5 AQI")
lines(test_monthly)

# Compare the forecast with the actual test data by calculating MAPE and MSE
# Mean Absolute Percentage Error (MAPE)
mape_snaive_monthly <- mean(abs((test_monthly - forecast_snaive_monthly$mean)/test_monthly))
mape_snaive_monthly

# Mean Absolute Error (MAE)
mae_snaive_monthly <- mean(abs((test_monthly - forecast_snaive_monthly$mean)))
mae_snaive_monthly

# Mean Squared Error (MSE)
mse_snaive_monthly <- mean((test_monthly - forecast_snaive_monthly$mean)^2)
mse_snaive_monthly

@@

#### Step 9b: Forecast using ETS - Monthly

```{r, tidy=TRUE}
# Build the exponential smoothing model
model_ets_monthly <- ets(train_monthly, lambda="auto")
summary(model_ets_monthly)
forecast_ets_monthly <- forecast(model_ets_monthly, h=31)

# Plot the forecasts for ETS model
plot(forecast_ets_monthly, main = "PM 2.5 AQI Forecast - ETS",
         xlab = "Month", ylab = "PM 2.5 AQI")
lines(test_monthly)

# Compare the forecast with the actual test data by calculating MAPE and MSE
# Mean Absolute Percentage Error (MAPE)
mape_ets_monthly <- mean(abs((test_monthly - forecast_ets_monthly$mean)/test_monthly))
mape_ets_monthly

# Mean Absolute Error (MAE)
mae_ets_monthly <- mean(abs((test_monthly - forecast_ets_monthly$mean)))
mae_ets_monthly

# Mean Squared Error (MSE)
mse_ets_monthly <- mean((test_monthly - forecast_ets_monthly$mean)^2)
mse_ets_monthly

# Evaluate the residuals
checkresiduals(forecast_ets_monthly)

Box.test(residuals(forecast_ets_monthly), lag = 12, type = "Ljung-Box")

#### Step 10b: Forecast using Neural Network - Monthly

model_nn_monthly <- nnetar(train_monthly,
                           P=12,
                           size=5, 
                           decay=0.5,
                           repeats=1000, 
                           lambda="auto")
summary(model_nn_monthly)
forecast_nn_monthly <- forecast(model_nn_monthly, PI=FALSE, h=31)

# Plot the forecasts for NN model
plot(forecast_nn_monthly, main = "PM 2.5 AQI Forecast - Neural Network",
         xlab = "Month", ylab = "PM 2.5 AQI")
lines(test_monthly)

# Compare the forecast with the actual test data by calculating MAPE and MSE
# Mean Absolute Percentage Error (MAPE)
mape_nn_monthly <- mean(abs((test_monthly - forecast_nn_monthly$mean)/test_monthly))
mape_nn_monthly

# Mean Absolute Error (MAE)
mae_nn_monthly <- mean(abs((test_monthly - forecast_nn_monthly$mean)))
mae_nn_monthly

# Mean Squared Error (MSE)
mse_nn_monthly <- mean((test_monthly - forecast_nn_monthly$mean)^2)
mse_nn_monthly

# Evaluate the residuals
checkresiduals(forecast_nn_monthly)
```
